\chapter{Introduction}

Precision medicine is a medical model that proposes the customization of
healthcare, with medicial decisions, practices, and/or products being tailored
to the individual patient. In this model, diagnostic testing is often
employed for selecting appropriate and optimal therapies based on the
context of a patient's genetic content or other molecular or cellular
analysis. Tools employed in precision medicine can include molecular
diagnostics, imaging, and analytics/software.

With new treatments and novel technology available, personalized medicine has
become an important piece in the new era of medical product
development. Traditional statistics methods for personalized medicine and
subgroup identification primarily focus on single treatment or two arm
randomized control trials. Motivated by the recent development of outcome
weighted learning framework, we propose an alternative algorithm to search
treatment assignments which has a connection with subgroup identification
problems. Our method focuses on applications from clinical trials to generate
easy to interpret results. This framework is able to handle two or more than two
treatments from both randomized control trials and observational studies.

Assume that there are $N$ subjects from a large population. For sample $i$,
$A_i$ is the treatment assignment, $Y_i$ is the response where larger value is
better, and $X_i$ is vector of covariates. Let $(Y, A, X)$ be the generic random
variable $\{(Y_i, A_i, X_i)\}$, ${\cal P}$ be the corresponding
distribution, and $E$ be the expectation with respect to ${\cal P}$. 
Let ${\cal D}(\cdot)$ be a treatment recommendation, i.e,
\[
  {\cal D(\cdot)}: \chi \to A, \quad X_i \in \chi,
\]
and ${\cal P^{\cal D}}$ be the distribution of $(Y, A, X)$ given $A = {\cal D}(X)$. 
Define
\[
E^{\cal D}(Y) = \int Y \, d{\cal P}^{\cal D} =
= \int Y \frac{d {\cal P}^{\cal D}}{d {\cal P}}
= E \left [
  \frac{I_{A = {\cal D}(X)}}{p(A|X)} Y
    \right], 
\]
where 
\[
\frac{d {\cal P}^{\cal D}}{d {\cal P}} =
\frac{p(y|x, a) I_{a = {\cal D}(x)} p(x)}{p(y|x, a) p(a|x)p(x)} =
\frac{I_{a = {\cal D}(x)}}{p(a|x)}
\]

The quantity $E^{{\cal D}}(Y)$ can be used to compare different treatment
recommendations. For instance, assume that we have two doctors, Adam and Barry,
each having a treatment rule. Doctor Adam gives patients treatment 1 if
$X \geq 2$ and treatment 0 otherwise, denoted by ${\cal D}_A$, and doctor Barry give
patients treatment 1 if $X \geq 3$ and treatment 0 otherwise, denoted by
${\cal D}_B$. Over the following population, we have

\begin{table}[htbp]
\centering
\caption{Comparison of treatment recommendations between Doctors Adam and Barry}
\begin{tabular}{c|ccc|ccccc}
\hline \hline
ID & $Y$ & $A$ & $X$ & $P\left( {A|X} \right)$ &
$\mathcal{D}_A$ & $\mathcal{D}_B$ & $\mathcal{D}_A = A$ & $\mathcal{D}_B = A$
\\
\hline
1  & 1 & 0 & 1 & 0.5  & 0  & 0  & 1   & 1   \\
2  & 2 & 0 & 2 & 0.5  & 1  & 0  & 0   & 1   \\
3  & 3 & 0 & 3 & 0.5  & 1  & 1  & 0   & 0   \\
4  & 4 & 0 & 4 & 0.5  & 1  & 1  & 0   & 0   \\
5  & 5 & 0 & 5 & 0.5  & 1  & 1  & 0   & 0   \\
6  & 3 & 1 & 1 & 0.5  & 0  & 0  & 0   & 0   \\
7  & 3 & 1 & 2 & 0.5  & 1  & 0  & 1   & 0   \\
8  & 3 & 1 & 3 & 0.5  & 1  & 1  & 1   & 1   \\
9  & 3 & 1 & 4 & 0.5  & 1  & 1  & 1   & 1   \\
10 & 3 & 1 & 5 & 0.5  & 1  & 1  & 1   & 1   \\
\hline \hline
\end{tabular}
\end{table}

\[
E^{{\cal D}_A} = \frac{1}{10}\times \frac{1}{0.5}
\left (
1\times 1 + 0 \times 2 + 0 \times 3 + 0 \times 4 + 0 \times 5 +
0\times 3 + 1 \times 3 + 1 \times 3 + 1 \times 3 + 1 \times 3
\right) = 2.6
\]
and
\[
E^{{\cal D}_B} = \frac{1}{10}\times \frac{1}{0.5}
\left (
1\times 1 + 1 \times 2 + 0 \times 3 + 0 \times 4 + 0 \times 5 +
0\times 3 + 0 \times 3 + 1 \times 3 + 1 \times 3 + 1 \times 3
\right) = 2.4
\]

Therefore, we conclude doctor Adam's treatment recommendation is better than
doctor Barry's treatment recommendation. When there are more than two treatment
rules, the objective is to find  ${\cal D}_o$ such that
\begin{equation}
  \label{eq:max}
  {\cal D}_o \in \mathop {\arg \max} \limits_{{\cal D} \in R} E^{\cal D}(Y) =
  \mathop {\arg \max} \limits_{{\cal D} \in R}
  E\left [ \frac{I_{A={\cal D}(X)}}{p(A|X)} Y \right ]
\end{equation}
where $R$ is the space of possible treatment recommendations. Alternatively, the
problem can be equivalently expressed as
\begin{equation}
  \label{eq:min}
  {\cal D}_o \in \mathop {\arg \min} \limits_{{\cal D} \in R} E
  \left [ \frac{I_{A \neq {\cal D}(X)}}{p(A|X)} Y \right ].
\end{equation}
